Uncovering underlying high-level musical content in the temporal domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn music embeddings with applications to MIR downstream tasks

### 2023-05-25 abstract/abstract.tex
Updates from Overleaf

Music, an integral part of human culture, offers a diverse yet complex field of study due to its intricate styles, mutable ground truth, and subjective nature. This research, conducted in collaboration with Epidemic Sound AB and the Music Technology Group (MTG) at Universitat Pompeu Fabra (UPF), employs self-supervised contrastive learning of musical representation to uncover the fundamental structure of Western tonal music.

Embracing a subtle novel approach, this process is facilitated by applying Siamese networks with a triplet loss, mimicking human aural skills in interpreting music to discern abstract and semantic musical elements, irrespective of the sonic qualities. The study replaces traditional acoustical features with deep audio embeddings to compute high-level, sound-agnostic, and content-sensitive music identification.

Results XXXXXX

The preliminary results suggest that our approach to learning music-informed embeddings holds significant potential for nearly all MIR downstream tasks. Music-motivated embeddings represent a promising technique, adaptable potentially to other tasks hindered by data scarcity. This method could significantly influence advancements in intelligent music recommendation systems and the efficient enforcement of intellectual property rights.

>>>>

This thesis asserts the existence of high-level musical concepts that, while evolving and unfolding over time, remain invariant to changes in sonic qualities. This notion echoes the symbolic domain's characteristics, wherein the essence persists regardless of varied interpretations through diverse performances, instruments, and styles.

In collaboration with Epidemic Sound AB and the Music Technology Group (MTG) at Universitat Pompeu Fabra (UPF), we employed self-supervised contrastive learning of musical representation to decipher the underlying structure of Western tonal music.

Taking a subtle yet novel approach, we utilized deep neural networks with a triplet loss function to facilitate this process. This strategy mirrors human aural skills in music interpretation, discerning abstract and semantic musical elements irrespective of their sonic qualities. This research supplants traditional acoustical features with deep audio embeddings, enabling high-level, sound-agnostic, and content-sensitive music identification.

Our method of learning embeddings is significant, emphasizing the use of full-resolution data to maintain the integrity of high-level musical information. Leveraging our expertise in the music domain, we construct aggressive transformations to embed heuristic musical concepts. With these innovative strides, we strive to bridge the gap between music and machine learning, improving the efficacy of machine listening models.

Preliminary results indicate that our musically-informed approach to learning embeddings carries the substantial potential for boundary detection tasks and, as our intuition suggests, possibly for nearly all Music Information Retrieval (MIR) downstream tasks. Music-motivated embeddings emerge as a promising technique, potentially adaptable to other tasks challenged by data scarcity.

### 2023-05-25 figures/neural networks/Multi-layer_NN.tex
Updates from Overleaf

\begin{figure}[ht]
	\centering
	\begin{tikzpicture}[shorten >=1pt]
		\tikzstyle{unit}=[draw,shape=circle,minimum size=1.15cm]
		%\tikzstyle{hidden}=[draw,shape=circle,fill=black!25,minimum size=1.15cm]
		\tikzstyle{hidden}=[draw,shape=circle,minimum size=1.15cm]

		\node[unit](x0) at (0,3.5){$x_0$};
		\node[unit](x1) at (0,2){$x_1$};
		\node at (0,1){\vdots};
		\node[unit](xd) at (0,0){$x_D$};

		\node[hidden](h10) at (3,4){$y_0^{(1)}$};
		\node[hidden](h11) at (3,2.5){$y_1^{(1)}$};
		\node at (3,1.5){\vdots};
		\node[hidden](h1m) at (3,-0.5){$y_{m^{(1)}}^{(1)}$};

		\node(h22) at (5,0){};
		\node(h21) at (5,2){};
		\node(h20) at (5,4){};

		\node(d3) at (6,0){$\ldots$};
		\node(d2) at (6,2){$\ldots$};
		\node(d1) at (6,4){$\ldots$};

		\node(hL12) at (7,0){};
		\node(hL11) at (7,2){};
		\node(hL10) at (7,4){};

		\node[hidden](hL0) at (9,4){$y_0^{(L)}$};
		\node[hidden](hL1) at (9,2.5){$y_1^{(L)}$};
		\node at (9,1.5){\vdots};
		\node[hidden](hLm) at (9,-0.5){$y_{m^{(L)}}^{(L)}$};

		\node[unit](y1) at (12,3.5){$y_1^{(L+1)}$};
		\node[unit](y2) at (12,2){$y_2^{(L+1)}$};
		\node at (12,1){\vdots};	
		\node[unit](yc) at (12,0){$y_C^{(L+1)}$};

		\draw[->] (x0) -- (h11);
		\draw[->] (x0) -- (h1m);

		\draw[->] (x1) -- (h11);
		\draw[->] (x1) -- (h1m);

		\draw[->] (xd) -- (h11);
		\draw[->] (xd) -- (h1m);

		\draw[->] (hL0) -- (y1);
		\draw[->] (hL0) -- (yc);
		\draw[->] (hL0) -- (y2);

		\draw[->] (hL1) -- (y1);
		\draw[->] (hL1) -- (yc);
		\draw[->] (hL1) -- (y2);

		\draw[->] (hLm) -- (y1);
		\draw[->] (hLm) -- (y2);
		\draw[->] (hLm) -- (yc);

		\draw[->,path fading=east] (h10) -- (h21);
		\draw[->,path fading=east] (h10) -- (h22);

		\draw[->,path fading=east] (h11) -- (h21);
		\draw[->,path fading=east] (h11) -- (h22);

		\draw[->,path fading=east] (h1m) -- (h21);
		\draw[->,path fading=east] (h1m) -- (h22);

		\draw[->,path fading=west] (hL10) -- (hL1);
		\draw[->,path fading=west] (hL11) -- (hL1);
		\draw[->,path fading=west] (hL12) -- (hL1);

		\draw[->,path fading=west] (hL10) -- (hLm);
		\draw[->,path fading=west] (hL11) -- (hLm);
		\draw[->,path fading=west] (hL12) -- (hLm);

		\draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt] (-0.5,4) -- (0.75,4) node [black,midway,yshift=+0.6cm]{input layer};
		\draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt] (2.5,4.5) -- (3.75,4.5) node [black,midway,yshift=+0.6cm]{$1^{\text{st}}$ hidden layer};
		\draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt] (8.5,4.5) -- (9.75,4.5) node [black,midway,yshift=+0.6cm]{$L^{\text{th}}$ hidden layer};
		\draw [decorate,decoration={brace,amplitude=10pt},xshift=-4pt,yshift=0pt] (11.5,4) -- (12.75,4) node [black,midway,yshift=+0.6cm]{output layer};
	\end{tikzpicture}
	\caption[Network graph for a $(L+1)$-layer perceptron.]{Network graph of a $(L+1)$-layer perceptron with $D$ input units and $C$ output units. The $l^{\text{th}}$ hidden layer contains $m^{(l)}$ hidden units.}
	\label{fig:multilayer-perceptron}
\end{figure}

>>>>

\begin{figure}[ht]
	\centering
	% NEURAL NETWORK with coefficients, shifted
\begin{tikzpicture}[x=2.2cm,y=1.4cm]
  \message{^^JNeural network, shifted}
  \readlist\Nnod{4,5,5,5,3} % array of number of nodes per layer
  \readlist\Nstr{n,m,m,m,k} % array of string number of nodes per layer
  \readlist\Cstr{\strut x,a^{(\prev)},a^{(\prev)},a^{(\prev)},y} % array of coefficient symbol per layer
  \def\yshift{0.5} % shift last node for dots

  \message{^^J  Layer}
  \foreachitem \N \in \Nnod{ % loop over layers
    \def\lay{\Ncnt} % alias of index of current layer
    \pgfmathsetmacro\prev{int(\Ncnt-1)} % number of previous layer
    \message{\lay,}
    \foreach \i [evaluate={\c=int(\i==\N); \y=\N/2-\i-\c*\yshift;
                 \index=(\i<\N?int(\i):"\Nstr[\lay]");
                 \x=\lay; \n=\nstyle;}] in {1,...,\N}{ % loop over nodes
      % NODES
      \node[node \n] (N\lay-\i) at (\x,\y) {$\Cstr[\lay]_{\index}$};

      % CONNECTIONS
      \ifnum\lay>1 % connect to previous layer
        \foreach \j in {1,...,\Nnod[\prev]}{ % loop over nodes in previous layer
          \draw[connect,white,line width=1.2] (N\prev-\j) -- (N\lay-\i);
          \draw[connect] (N\prev-\j) -- (N\lay-\i);
          %\draw[connect] (N\prev-\j.0) -- (N\lay-\i.180); % connect to left
        }
      \fi % else: nothing to connect first layer

    }
    \path (N\lay-\N) --++ (0,1+\yshift) node[midway,scale=1.5] {$\vdots$};
  }

  % LABELS
  \node[above=0.8,align=center,mygreen!60!black] at (N1-1.90) {input\\[-0.2em]layer};
  \node[above=0.5,align=center,myblue!60!black] at (N3-1.90) {hidden layers};
  \node[above=1.3,align=center,myred!60!black] at (N\Nnodlen-1.90) {output\\[-0.2em]layer};

\end{tikzpicture}
	\caption[Network graph for perceptron.]{\small{Network graph of a perceptron with $D$ input units and $C$ output units. The $l^{\text{th}}$ hidden layer contains $m^{(l)}$ hidden units. Each neuron in a layer receives input from the previous layer and computes an output value using an activation function. The output of the last layer represents the prediction or classification result.
}}
	\label{fig: multilayer color perceptron}
\end{figure}

### 2023-05-25 figures/neural networks/activation.tex
Updates from Overleaf

% NEURAL NETWORK activation
% https://www.youtube.com/watch?v=aircAruvnKk&list=PLZHQObOWTQDNU6R1_67000Dx_ZCJB-3pi&index=1
\begin{tikzpicture}[x=2.7cm,y=1.6cm]
  \message{^^JNeural network activation}
  \def\NI{5} % number of nodes in input layers
  \def\NO{4} % number of nodes in output layers
  \def\yshift{0.4} % shift last node for dots

@@

\end{tikzpicture}

>>>>

\begin{figure}
    \centering
    \scalebox{0.9}{
% NEURAL NETWORK activation
% https://www.youtube.com/watch?v=aircAruvnKk&list=PLZHQObOWTQDNU6R1_67000Dx_ZCJB-3pi&index=1
\begin{tikzpicture}[x=2.7cm,y=1.6cm]
\centering
  \message{^^JNeural network activation}
  \def\NI{5} % number of nodes in input layers
  \def\NO{4} % number of nodes in output layers
  \def\yshift{0.4} % shift last node for dots

@@

\end{tikzpicture}
}
\caption[Activation function.]{\small{The input layer with \textit{n} nodes is denoted as $a_i^{(0)}$ for $i \in \{1, \ldots, n\}$. The output layer with \textit{m} nodes is $a_j^{(1)}$ for $j \in \{1, \ldots, m\}$. Connections between these layers are governed by the weight matrix $\mathbf{W}^{(0)} \in \mathbb{R}^{m\times n}$ and bias vector $\mathbf{b}^{(0)} \in \mathbb{R}^{m}$, with elements $w_{j,i}$ and $b_j^{(0)}$ respectively. The activation function, denoted as $\sigma$, operates element-wise on the sum of the product of the weight matrix, input vector, and bias vector.}}

    \label{fig:activation_function}

\end{figure}

### 2023-05-25 methods/subsections/machine learning.tex
Updates from Overleaf

Deep learning (DL) is a subfield of machine learning that employs deep neural networks to learn hierarchical data representations, enabling computational models with multiple processing layers to learn data abstractions across numerous levels. These methods have dramatically advanced the state-of-the-art in various domains, including speech recognition, visual object recognition, object detection, drug discovery, genomics, and more \cite{LeCun2015DeepLearning}. The effectiveness of deep learning arises from its nonlinear modeling capabilities and scalability with large datasets.

A central algorithm in deep learning is backpropagation, which uses the forward and backward pass process to compute gradients of the loss function concerning the network's parameters. The forward pass involves propagating input data through the network to generate predictions. In contrast, the backward pass calculates the error between predictions and ground truth and returns it through the network. This procedure updates the network's parameters minimizing the loss function and enhancing the model's performance.

@@

\subsection{Artificial Neural Networks (ANNs)}
Artificial Neural Networks (ANNs) are computational models inspired by biological neural systems. They use interconnected layers of artificial neurons to process information and perform tasks like pattern recognition, classification, and regression. ANNs adjust the connections or weights between neurons during learning to improve performance. The concept of neural networks emerged in the mid-20th century with the work of Warren McCulloch and Walter Pitts. Still, modern neural networks began with Frank Rosenblatt's perceptron in 1957, a single-layer network able to learn linearly separable patterns.

%Single neuron
\input{figures/neural networks/Single Neuron}

%NN
\input{figures/neural networks/Network_graph}

It consists of an input layer, one or more hidden layers, and an output layer. Each neuron in a layer receives input from the previous layer and computes an output value using an activation function. The output of the last layer represents the prediction or classification result.

@@

>>>>

Deep learning (DL), a subset of machine learning, employs deep neural networks to decipher hierarchical data representations, allowing models with multiple processing layers to abstract data across numerous levels \cite{LeCun2015DeepLearning}. DL models have significantly improved various domains due to their nonlinear modeling capabilities and scalability with large datasets.

They leverage the central algorithm, backpropagation, which calculates gradients of the loss function related to network parameters through forward and backward pass processes, thereby optimizing model performance.

@@

Deep Neural Networks (DNNs), computational models rooted in biological neural systems, utilize layers of artificial neurons for tasks such as pattern recognition, classification, and regression. DNNs evolve by adjusting connections between neurons during learning, a concept initiated in the mid-20th century but modernized with Frank Rosenblatt's single-layer perceptron in 1957, capable of learning linearly separable patterns.

Another crucial component in the operation of DNNs is the activation function. This function is applied to a neuron's inputs, transforming the input's summed product and weights into an output passed onto the next layer. It introduces non-linearity into the network, enabling it to learn complex patterns and relationships from the data.

@@

%Colour CNN
\input{figures/neural networks/colour CNN}

### 2023-05-25 methods/subsections/model architecture and training strategy.tex
Updates from Overleaf

The original model has been modified slightly to serve our specific needs: the output from the final convolutional layer is now subjected to an average pooling operation across the temporal dimension.

@@

The maximum array length in the batch is given by:

@@

This adaptation of the original architecture incorporates an average pooling step at the end.

Representation of the Average Pooling operation. Here, $y_i$ is the output tensor after the average pooling process. Each element $x_{ij}$ is part of the 2D input tensor. $N$ is the count of elements in the second dimension of the input tensor, while $i$ and $j$ are used to iterate over the first and second dimensions of the tensors, respectively.

\input{figures/neural networks/dropout}

\input{figures/my model/SampleCNN}

%Colour CNN
\input{figures/neural networks/colour CNN}

@@

>>>>

The original model has been modified slightly to serve our specific needs: the output from the final convolutional layer is now subjected to an average pooling operation at the end.

Here, $y_i$ is the output tensor after the average pooling process. Each element $x_{ij}$ is part of the input tensor. $N$ is the count of elements in the second dimension of the input tensor, while $i$ and $j$ are used to iterate over the first and second dimensions of the tensors, respectively.

@@

In implementing batch normalization, it is necessary to standardize the audio lengths across all elements in the minibatch. We opted to zero-pad all clips to the length of the longest clip, valuing data integrity and completeness over potential performance trade-offs. Thus, the length of the longest array in the batch, which sets the standard for all others, is as follows:

@@

@@

\begin{figure}
    \centering
    \includegraphics[clip,width=\columnwidth]{figures/images/Triplet Network.jpeg}
    \caption[]{Caption}
    \label{fig:my_label}
\end{figure}
